A Short and Mechanized Consistency Proof for Dependent Type Theory

### 2023-12-31 paper/paper.tex
Minor

@@

\begin{lemma}[ST-Pi]
  If $[[G |= A : Set i]]$ and $[[G ++ A |= B : Set i]]$, then $[[G |= Pi
  A B : Set i]]$.
\end{lemma}
\begin{proof}
  Applying Lemma~\ref{lemma:setinv} to the
  conclusion, it now suffices to show that given $[[rho |= G]]$, there
  exists some $[[S]]$ such that $[[InterpR i Pi A B S]]$.
  From Lemma~\ref{lemma:setinv} and $[[G |= A : Set i]]$, we know that
  there exists some set $[[S0]]$ such that $[[InterpR i A S0]]$.
  Let $[[F]]$ be the following relation:
  \[ (a, S) \in [[F]] \iff [[InterpR i B {rho .: a} S]] \]
  By Lemma~\ref{lemma:logreldeter}, $[[F]]$ is a partial
function. From $[[G ++ A |= B : Set i]]$, we know that there must
exists $[[S]]$ such that $[[InterpR i B {rho .: a} S]]$ for every $[[a
in S0]]$. As a result, $[[F]]$ is defined on every element of $[[S0]]$.
We can now instantiate  $[[S]]$ with the set $[[ { b | forall a, (# a
  in S0 implies b a in F a #) }]]$. The goal $[[InterpR i Pi A B S]]$
 immediately follows from  \rref{I-Pi} and the definition of $[[F]]$.
\end{proof}

>>>>

The semantic context well-formedness judgment ($[[|= G]]$), unlike its syntactic
counterpart $[[|- G]]$, is defined through a for all quantified statement rather
than inductively over the context. It is easy to recover the same
structural rules:
\begin{lemma}[TODO]

Likewise, the semantic well-formendness judgment for substitutions
satisfies similar structural rules.
\begin{lemma}[Well-formed $[[rho]]$ cons]
  If $[[InterpR i A S]]$, $[[a in S]]$, and $[[rho |= G]]$, then
  $[[rho .: a |= G ++ A]]$.
\end{lemma}

@@

\begin{lemma}[ST-Pi]
  If $[[G |= A : Set i]]$ and $[[G ++ A |= B : Set i]]$, then $[[G |= Pi
  A B : Set i]]$.
\end{lemma}
\begin{proof}
  Applying Lemma~\ref{lemma:setinv} to the
  conclusion, it now suffices to show that given $[[rho |= G]]$, there
  exists some $[[S]]$ such that $[[InterpR i Pi A B S]]$.
  From Lemma~\ref{lemma:setinv} and $[[G |= A : Set i]]$, we know that
  there exists some set $[[S0]]$ such that $[[InterpR i A S0]]$.
From $[[G ++ A |= B : Set i]]$, we know that there must
exists $[[S]]$ such that $[[InterpR i B {rho .: a} S]]$ for every $[[a
in S0]]$. The conclusion immediately follows from Lemma~\ref{lemma:piintroalt}.
\end{proof}
